---
title: "p8105_hw3_md4270"
author: "Manye Dong"
date: "2023-10-07"
output: github_document
---
```{r message=FALSE}
library(tidyverse)
```

## Problem 1
```{r}
library(p8105.datasets)
data("instacart")
```

The `instacart` dataset has `r nrow(instacart)` number of rows and `r ncol(instacart)` number of columns. Each row is a single product from an instacart order placed. Key variables contain the order id, product id, user id, the order date, product id and the aisle it's in.

In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

```{r message=FALSE}
aisle_num = length(unique(instacart$aisle_id))
```
```{r message=FALSE}
aisle_most = instacart |>
  group_by(aisle) |>
  summarise(num_products = n()) |>
  arrange(desc(num_products))

head(aisle_most)
```

There are `r aisle_num` number of aisles, and `r aisle_most$aisle[1]` is the aisle where most items are ordered from, `r aisle_most$aisle[2]` following it.

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r message=FALSE, warning=FALSE}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_bar(stat="identity", fill = "skyblue", color = "black") +
  labs(title = "Number of Items Ordered In Each Aisle", 
       x = "Aisle", 
       y = "Number of Items") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r message=FALSE, warning=FALSE}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable(title = "Top 3 Most Popular Items in Selected Aisles", 
               col.names = c("Aisle", "Product Name", "Numbers Sold", "Rank"))
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r message=FALSE, warning=FALSE}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarise(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2, 
               col.names = c("Product Name", "Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"))
```


## Problem 2
```{r}
library(p8105.datasets)
data("brfss_smart2010")
```
```{r message=FALSE, warning=FALSE}
head(brfss_smart2010)
```
Clean the data based on instructions:
```{r, message=FALSE, warning=FALSE}
brfss_smart2010_cleaned = 
  brfss_smart2010 |>
  janitor::clean_names() |>
  rename("state"="locationabbr", "county"="locationdesc") |> 
  filter(topic=="Overall Health" & response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) |>
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))
  
head(brfss_smart2010_cleaned)
```

```{r message=FALSE, warning=FALSE}
locations_2002 = 
  brfss_smart2010_cleaned |>
  filter(year==2002) |>
  group_by(year, state) |>
  summarise(num_location = n_distinct(county)) |>
  filter(num_location >= 7)
```
In 2002, states `r locations_2002$state` were observed at 7 or more locations. 

```{r message=FALSE, warning=FALSE}
locations_2010 = 
  brfss_smart2010_cleaned |>
  filter(year==2010) |>
  group_by(year, state) |>
  summarise(num_location = n_distinct(county)) |>
  filter(num_location >= 7)
```
In 2010, states `r locations_2002$state` were observed at 7 or more locations. 

Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state: 
```{r message=FALSE, warning=FALSE}
avg_values = 
brfss_smart2010_cleaned |>
  filter(response=="Excellent") |>
  select(-county) |>
  group_by(year, state) |>
  mutate(avg_data_value = mean(data_value, na.rm = TRUE)) |>
  select(year, state, avg_data_value)
  
head(avg_values)
```

Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
```{r message=FALSE, warning=FALSE}
ggplot(avg_values, aes(x = year, y = avg_data_value, color = state, group = state)) +
  geom_line() +
  labs(x = "Year", y = "Average Value", title = "Spaghetti Plot of Average Values Over Time by State") +
  theme_minimal()
```

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State:
```{r message=FALSE, warning=FALSE}
ny_data = 
  brfss_smart2010_cleaned |>
  filter(brfss_smart2010_cleaned$state == "NY" 
       & brfss_smart2010_cleaned$year %in% c(2006, 2010))
```
```{r message=FALSE, warning=FALSE}
ggplot(ny_data, aes(x = response, y = data_value, fill = response)) +
  geom_boxplot() +
  facet_wrap(~year, scales = "free_y") +
  scale_fill_brewer(palette = "Pastel1", name = "Response") +
  labs(title = "Distribution of Data Value for Responses in NY State",
       y = "Data Value",
       x = "Response") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


## Problem 3
Clean the data according to instructions:
```{r message=FALSE}
library(readr)
nhanes_accel = read_csv("nhanes_accel.csv")
nhanes_covar = read_csv("nhanes_covar.csv", skip = 4)
```

```{r message=FALSE, warning=FALSE}
nhanes_accel_cleaned = 
  nhanes_accel |>
  janitor::clean_names()
```

```{r message=FALSE, warning=FALSE}
nhanes_covar_cleaned =
  nhanes_covar |>
  janitor::clean_names() |>
  mutate(
      sex = recode(sex, "1" = "male", "2" = "female"),
      education = recode(education, "1" = "Less than high school", 
                         "2" = "High school equivalent", 
                         "3" = "More than high school")) |>
  mutate(sex = factor(sex), 
         education = factor(education)) |>
  filter(age >= 21) |>
  drop_na()
```
Now, we join the two cleaned tables to include all valid observations:
```{r message=FALSE, warning=FALSE}
nhanes = merge(nhanes_covar_cleaned, nhanes_accel_cleaned, by = "seqn")
```
Produce a reader-friendly table for the number of men and women in each education category:
```{r message=FALSE, warning=FALSE}
nhanes_sex_tbl = 
  nhanes |>
  group_by(education, sex) |>
  summarise(num_people = n_distinct(seqn))|>
  pivot_wider(names_from = sex, values_from = num_people) |>
  arrange(education) |>
  knitr::kable(title = "Number of Men and Women in Each Education Category", 
               col.names = c("Education", "Female", "Male"))

nhanes_sex_tbl
```
Next, create a visualization of the age distributions for men and women in each education category:
```{r message=FALSE, warning=FALSE}
ggplot(nhanes, aes(x = education, y = age, fill = sex)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Pastel1", name = "Sex") + 
  labs(title = "Age Distribution by Gender and Education",
       x = "Education Level",
       y = "Age") +
  theme_minimal()
```

Based on the table above, there are more female in the higher education group, while fewer female in the high school equivalent background compared to male.

The boxplot that displays the age distribution of participants across different education levels. 

Among those who participated in the NHANES study, the group with "high school or equivalent" degree has the highest median age for females. The group with "more than high school degree" has the lowest median age for females.

Among those who participated in the NHANES study, the group with "less than high school" degree has the highest median age for males. The group with "more than high school degree" has the lowest median age for males.

Using the tidied dataset from above, aggregate across minutes to create a total activity variable for each participant:
```{r message=FALSE, warning=FALSE}
total_act = 
  nhanes |>
  mutate(total_activity = rowSums(across(min1:min1440), na.rm = TRUE))
```
Make a 3-panel plot:
```{r message=FALSE, warning=FALSE}
ggplot(total_act, aes(x = age, y = total_activity, color = sex)) +
  geom_point(alpha = .5) +
  facet_wrap(~education) +
  geom_smooth() +
  labs(title = "Total Activity for Male vs Female", 
       x = "Age", 
       y = "Total Activity")
```

For all groups of participants, the general trend of total activity time versus age is decreasing, which makes sense because elder people tend to do less exercise.

For "high school equivalent" and "more than high school" groups, male generally have fewer total activity time across each age bins. Total activity time plot for participants with "high school equivalent" degree shows a "peak" of activity time at around age 40 for both genders. For the other two groups, the "peak" is around at 60.

Lastly, we make a 24-hour activity time courses plot:
```{r message=FALSE, warning=FALSE}
total_act_24hrs = 
  nhanes |>
  group_by(education, sex) |>
  summarise(across(starts_with("min"), ~ mean(.), .names = "mean_{.col}")) |>
  pivot_longer(
    cols = starts_with("mean_"), 
    names_to = "time", 
    values_to = "mean") |>
  mutate(time = substring(time, 9),
         time = as.numeric(time))
```
Make a 3-panel plot:
```{r message=FALSE, warning=FALSE}
ggplot(total_act_24hrs, aes(x = time, y = mean, color = sex)) +
  geom_line() +
  facet_grid(. ~education) +
  labs(title = "24-Hour Activity by Education and Gender", 
       x = "Time (min)", 
       y = "Activity")
```

All three plots show a surprisingly similar trends for 24-hour activity time change. For the group with "less than high school" degree, there is not much difference across male and female (two lines basically overlap). For the other two, female tend to have a higher mean value of activity time than male do.