---
title: "p8105_hw3_md4270"
author: "Manye Dong"
date: "2023-10-07"
output: github_document
---
```{r message=FALSE}
library(tidyverse)
```

## Problem 1
```{r}
library(p8105.datasets)
data("instacart")
```

```{r}
head(instacart)
```
The `instacart` dataset has `r nrow(instacart)` number of rows and `r ncol(instacart)` number of columns. Each row is an order placed. Key variables contain the order id, product id, user id, the order date, product id and the aisle it's in.
[giving illustrative examples of observations.]

```{r}
aisle_num = length(unique(instacart$aisle_id))
```
```{r}
aisle_most = instacart |>
  group_by(aisle) |>
  summarise(num_products = n()) |>
  arrange(desc(num_products))

aisle_most
```

* There are `r aisle_num` number of aisles, and `r aisle_most$aisle[1]` is the aisle where most items are ordered from.

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
instacart |>
  group_by(aisle) |>
  summarize(items_num = n()) |>
  filter(items_num > 10000) |>
  arrange(desc(items_num)) |>
  ggplot(aes(x=aisle, y=items_num)) +
  geom_bar(stat="identity", fill = "skyblue", color = "black") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Aisle", y = "Number of Items", title = "Items in 39 Aisles")
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r message=FALSE, warning=FALSE}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```


## Problem 2
```{r}
library(p8105.datasets)
data("brfss_smart2010")
```
```{r}
head(brfss_smart2010)
```
format the data to use appropriate variable names;
focus on the “Overall Health” topic
include only responses from “Excellent” to “Poor”
organize responses as a factor taking levels ordered from “Poor” to “Excellent”
```{r}
brfss_smart2010_cleaned = 
  brfss_smart2010 |>
  janitor::clean_names() |>
  rename("state"="locationabbr", "county"="locationdesc") |> 
  filter(topic=="Overall Health" & response %in% c("Excellent", "Very good", "Good", "Fair", "Poor"))
  
brfss_smart2010_cleaned$response = 
  factor(brfss_smart2010_cleaned$response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"), ordered = TRUE)
  
brfss_smart2010_cleaned
```

```{r}
locations_2002 = 
  brfss_smart2010_cleaned |>
  filter(year==2002) |>
  group_by(state) |>
  summarize(num_location = n_distinct(county)) |>
  filter(num_location >= 7)

head(locations_2002)
```
In 2002, state in the list `r locations_2002$state` were observed at 7 or more locations. 

```{r}
locations_2010 = 
  brfss_smart2010_cleaned |>
  filter(year==2010) |>
  group_by(state) |>
  summarize(num_location = n_distinct(county)) |>
  filter(num_location >= 7)

locations_2010
```
In 2010, state `r locations_2002$state` were observed at 7 or more locations. 

Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 
```{r}
avg_values = 
brfss_smart2010_cleaned |>
  filter(response=="Excellent") |>
  select(-county) |>
  group_by(year, state) |>
  mutate(avg_data_value = mean(data_value, na.rm = TRUE)) |>
  select(year, state, avg_data_value)
  
avg_values
```

Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
```{r}
ggplot(avg_values, aes(x = year, y = avg_data_value, color = state, group = state)) +
  geom_line() +
  labs(x = "Year", y = "Average Value", title = "Spaghetti Plot of Average Values Over Time by State") +
  theme_minimal()
```

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
ny_data = brfss_smart2010_cleaned |>
filter(brfss_smart2010_cleaned$state == "NY" 
       & brfss_smart2010_cleaned$year %in% c(2006, 2010))
```
```{r}
ggplot(ny_data, aes(x = response, y = data_value, fill = response)) +
  geom_boxplot() +
  facet_wrap(~year, scales = "free_y") +
  labs(title = "Distribution of data_value for Responses in NY State",
       y = "Data Value",
       x = "Response") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


## Problem 3
```{r}

```

